Pillar 1 - Sequential Pattern Mining

### 2019-01-23 Pillar 1 - Sequential Pattern Mining.Rmd
computes statistics (Conf_corr) and beggins to add plots

library(rlang)
```

@@

#Getting graphics
```{r}

@@

category <-quo(Discrete_grade) #Outcome or Low_grade or Discrete_grade or #None

@@

    mutate(
      Grade = case_when(
        Grade == "NG"~ "-1",
        T            ~ Grade),
      Grade = str_replace(string = Grade, pattern = ",",replace = "."),
      Grade = as.numeric(Grade)) %>%

@@

frequent_trajectories <- cspade(transactions, parameter = list(support = 0.01), control   = list(verbose = TRUE))

summary(frequent_trajectories)

rules <- ruleInduction(frequent_trajectories, confidence = 0.1, #we need to play with the parameters.
                    control    = list(verbose = TRUE))

@@

  filter(str_detect(RHS, string_filter)) %>% #FILTER (YES) --> uncomment, (NO)-->comment

  arrange(desc(confidence), desc(lift))

pass_rules_df

@@

>>>>

library(rlang)

library(ggplot2)
```

@@

@@

category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None

@@

    # mutate(
    #   Grade = case_when(
    #     Grade == "NG"~ "-1",
    #     T            ~ Grade),
    #   Grade = str_replace(string = Grade, pattern = ",",replace = "."),
    #   Grade = as.numeric(Grade)) %>% #this was already done in data preparation

@@

frequent_trajectories <- cspade(transactions, parameter = list(support = 0,
                                                               maxsize = 1,
                                                               maxlen = 2), control   = list(verbose = TRUE))

summary(frequent_trajectories)
as(frequent_trajectories, "data.frame")

rules <- ruleInduction(frequent_trajectories,
                       confidence = 0, #we need to play with the parameters.
                       control    = list(verbose = TRUE))

@@

#  filter(str_detect(RHS, string_filter)) %>% #FILTER (YES) --> uncomment, (NO)-->comment

  arrange(
    LHS, RHS,
    desc(confidence), desc(lift))

n_students <- length(unique(d_sequence_transactions$`Student ID`))

manual_confidence_rules <- pass_rules_df %>%
  separate(LHS, into=c("LHS Course ID", "LHS Score"), sep="_", remove= FALSE) %>%
  separate(RHS, into=c("RHS Course ID", "RHS Score"), sep="_", remove= FALSE) %>%
  mutate(`RHS Course ID`= substr(`RHS Course ID`, start=4, stop=10))%>%
  left_join(lookup_table, by = c("RHS Course ID"="Course ID")) %>%

  # Compute statistics
  group_by(LHS,`RHS Course ID`) %>%
  mutate(Prob_both = sum(support), 
         Conf_corr = support/Prob_both,
         n_rule = support * n_students,
         n_both =Prob_both * n_students,
         lift_diff= Conf_corr - Prop_low,
         lift_corr=Conf_corr / Prop_low) %>%
  ungroup() %>%

  # Filter and edit output
  filter(str_detect(`RHS Score`, pattern = "low"),
         str_detect(`LHS Score`, pattern = "low"),
         n_rule >= 5,
         Conf_corr > 0.2) %>%
  select(LHS, RHS, support, n_rule, Prob_both, n_both, Prop_low, Conf_corr, lift_corr, lift_diff) %>%
  mutate(support = round(support,3),
         Prob_both =round(Prob_both,3),
         Prop_low =round(Prop_low,3),
         Conf_corr= round(Conf_corr,3),
         lift_diff= round(lift_diff,3),
         lift_corr= round(lift_corr,3)) %>%
  arrange(lift_diff)

#FOR PLOTTING
for_plots <-  manual_confidence_rules %>%
  group_by(LHS)%>%
  mutate(LHS_count=n())

```
#Visualizing rules
```{r}
#look at counts
ggplot(for_plots, aes(LHS))+
  geom_bar()

#jitter
ggplot(manual_confidence_rules)+
  geom_jitter(mapping=aes(x=Conf_corr, y=log(support), color=lift_corr))

ggplot(manual_confidence_rules)+
  geom_jitter(mapping=aes(x=log(support), y=lift_corr, color = Conf_corr))

#heatmap
ggplot(select(manual_confidence_rules,LHS,RHS, n_rule), aes(x=RHS,y=LHS, fill=n_rule))+
  geom_raster() +
  theme(
    axis.text.x = element_text(angle=90)
  )

gglot(table(select(manual_confidence_rules,LHS,RHS)))

hi <- table(select(manual_confidence_rules,LHS,RHS))
#ggnetwork

@@

We want to change the calculation of confidence and support. The new rules are:
confidence: number of transactions of RHS/ number of transactions of RHS and compliment= p(A^B)/P(BUBc)
support:number of transactions of right hand side/number of transactions of RHS and compliment= p(A^B)/p(AUBUBc)
lift:

#Experimenting with transactions
```{r}
fail_threshold <- 6.5

category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None

lookup_table <- d_course %>%

  left_join(d_transcript, by = "Course ID") %>%
  mutate(
    Outcome = 
           case_when( Grade == -1   ~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade  > 5.5 ~"Pass",
                               T~"Fail"),
    Low_grade =
           case_when(Grade == -1 ~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                               T~"low score"),
    Discrete_grade =
           case_when(Grade <= 5.4             ~ 5.0,
                     between(Grade, 5.5, 6.4) ~ 6.0,
                     between(Grade, 6.5, 7.4) ~ 7.0,
                     between(Grade, 7.5, 8.4) ~ 8.0,
                     between(Grade, 8.5, 9.4) ~ 9.0,
                     Grade >= 9.5             ~ 10.0,
                     T ~ -99999    #this is a control, we should not have any grades here
                     ),
    None= "" # this might not be optimal
         ) %>%

  #mutate(Item = paste(`Course ID`, !!category, sep = "_")) %>% #Change for Outcome/Low_grade/Discrete_grade

  select(`Student ID`, #Item, 
         `Course ID`, !!category, Time) %>% # Course ID and Time may not be necessary

  group_by(`Course ID`, !!category) %>% ##this makes no sense we should just group id and low_grade

  summarize(Count = n()) %>%
  #separate(Item, into= c("Course ID", quo_name(category)), sep="_", remove = F) %>%
  ungroup() %>%
  group_by(`Course ID`) %>%
  mutate(Total= sum(Count))%>%
  filter(Low_grade=="low score")%>%
  mutate(Prop_low= Count/Total)

lookup_table
```

```{r}
pass_rules_df
test <- pass_rules_df %>%
  separate(LHS, into=c("c1","c2", "c3", "c4", "c5", "c6","c7", "c8", "c9", "c10"), sep= ",", remove=F) #has values until c6

```{r}

length(unique(d_transcript$`Student ID`))
```
